Progress Report for CSS 586: \\ Modeling Latent Patterns in Music

### 2021-04-19 papers/progress/alex/progress-alex.tex
save report updates

MusicVAE is a variational LSTM autoencoder for MIDI that incorporates a
novel hierarchical structure using a "composer" recurrent layer in its encoder
model to better capture structure at multiple
levels \cite{roberts_hierarchical_2018}.

@@

Other research has focused on modeling raw audio waveforms directly. WaveNet is
a causal convolutional neural network for generating raw audio waveforms,
developed by Google DeepMind, which achieves state of the art performance in
generating natural sounding speech from text, but is also capable of generating
short, realistic snippets of audio music \cite{oord_wavenet_2016}. Another model
named SampleRNN generates raw audio waveforms using a three-tier hierarchy of
gated recurrent units (GRU) to model recurrent structure at multiple temporal
resolutions \cite{mehri_samplernn_2017}.

A paper from Boston University describes an effort to combine the symbolic and
waveform approaches to music modeling \cite{manzelli_conditioning_2018}.

>>>>

MusicVAE is a variational LSTM autoencoder for MIDI that incorporates a
novel hierarchical structure using a ``composer'' recurrent layer in its encoder
model to better capture structure at multiple
levels \cite{roberts_hierarchical_2018}.

@@

The major advantage of working with the symbolic representation of music is that
it is of far lower dimensionality than the raw audio waveforms of a recorded
performance,

Other research has focused on modeling raw audio waveforms directly. WaveNet is
a causal convolutional neural network for generating raw audio waveforms,
developed by Google DeepMind, which achieves state of the art performance in
generating natural sounding speech from text, but is also capable of generating
short, realistic snippets of audio music \cite{oord_wavenet_2016}.

Another model named SampleRNN generates raw audio waveforms using a three-tier
hierarchy of gated recurrent units (GRU) to model recurrent structure at
multiple temporal resolutions \cite{mehri_samplernn_2017}.

Jukebox by OpenAI utilizes a vector-quantized variational autoencoder (VQ-VAE)
to compress raw audio into a sequence of discrete codes and models these
sequences using autoregressive transformers to generate music
\cite{dhariwal2020jukebox}.

Audio data can also be modeled in the frequency domain through the use of
Fourier analysis. The recent MelNet model is trained on spectrograms and can
learn musical structures such as melody and harmony and variations in volume,
timbre and rhythm \cite{vasquez2019melnet}.

A paper from Boston University describes an effort to combine the symbolic and
waveform approaches to music modeling, by training an LSTM to learn melodic
structure of different styles of music, then providing generations from this
model as conditioning inputs to a WaveNet-based raw audio generator
\cite{manzelli_conditioning_2018}.
